---
title: datalang
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(datalang)
my_spec <- system.file("specs/thisweek.yml", package = "datalang")
esta_semana <- translate_data(my_spec)


```

[![Travis build status](https://travis-ci.org/edgararuiz/datalang.svg?branch=master)](https://travis-ci.org/edgararuiz/datalang)
[![Coverage status](https://codecov.io/gh/edgararuiz/datalang/branch/master/graph/badge.svg)](https://codecov.io/github/edgararuiz/datalang?branch=master)
[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/datalang)](http://cran.r-project.org/package=datalang)

The goal is to help people whose first, or preferred, language is not English.  This package aids in translating the data and help files.  The hope is to lower the barrier of learning R, and also data analysis.

Even though this package can translate data on the fly, this package is meant to be used inside another R package.

## Installation

The development version is availble on [GitHub](https://github.com/):

``` r
# install.packages("devtools")
devtools::install_github("edgararuiz/datalang")
```

## How it works

The `datalang` package uses a translation spec file for translation.  The expected file format is YAML. `datalang` uses the YAML file to know which column names, and data values, to change.  The same YAML spec file can contain the information needed to produce a help file.

Here is a sample spec file:

```yaml
df:
  source: thisweek
  name: esta_semana
variables:
  day:
    trans: dia
    desc: Dia de la semana
    values:
      friday: viernes
      saturday: sábado
      sunday: domingo
  morning:
    trans: manana
    desc: Temperatura a las 10 AM
  afternoon:
    trans: tarde
    desc: Temperatura a las 3 PM
help:
  name: esta_semana
  alias: esta_semana
  title: Un data set de ejemplo
  description: Tres días de temperaturas tomadas en la mañana y la tarde
```

So a the data set can go from this:

```{r, echo = FALSE}
datalang::thisweek
```

To this:

```{r, echo = FALSE}
esta_semana
```

## Basics

The `datalang` package includes a simple data set called: `thisweek`.  The package also has a YAML spec file with the translation. 

```{r}
my_spec <- system.file("specs/thisweek.yml", package = "datalang")
```

The `translate_data()` function can ready the YAML spec file found in the `my_spec` file path. It uses that spec to locate the data set, in this case `thisweek`, modifies a copy of the data set based on the spec, and outputs a new, translated, data set. 

```{r}
translate_data(my_spec)
```

Not all of the spec's entries were needed to translate the data.  The other entries are useful for creating the help documentation.  In `datalang` there are two functions that output the help:

1. `create_rd()` - It outputs a character vector with contents that can be saved as an rd file.
1. `create_html_help()` - It outputs a character vector with contents that can be saved as an HTML file.


